---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### The data science workflow

<img src=https://i.imgur.com/MitpHrJ.png width=700>

https://r4ds.had.co.nz/introduction.html

### The data science toolkit

<img src=https://i.imgur.com/i61n3xb.png width=700>

https://www.tidyverse.org/

```{r packages}
library(tidyverse)
```

### Visualize: The ggplot template

<img src=https://i.imgur.com/WsoXgV2.png width=700>

Example:

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

https://ggplot2.tidyverse.org/

Template:

```
    ggplot(data = <DATA>) +
      <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

### Transform: Key dplyr verbs and how to compose them

Example:

```{r}
diamonds
```

```{r}
select(diamonds, cut, price)
```

Same using the "pipe" (`%>%` ~ "then"). Read it like this:

* Take the dataset `diamonds` **then**
* `select` the columns `cut` and `price`

```{r}
diamonds %>% 
  select(cut, price)
```

Another example:

* Take the dataset `diamonds` **then**
* `select` the columns `cut` and `price` **then**
* `count` unique values of the column `cut` **then**
* `filter` rows where `n` is greater than `1000`

```{r}
diamonds %>% 
  select(cut, price) %>% 
  count(cut) %>% 
  filter(n > 10000)
```

The traditional "nested" syntax produces the same output but is less readable:

```{r}
filter(count(select(diamonds, cut, price), cut), n > 10000)
```

https://dplyr.tidyverse.org/

### Tidy data: What and why

> Tidy datasets are easy to manipulate, model and visualise, and have a specific structure: each variable is a column, each observation is a row.

<img src=https://i.imgur.com/nBC5Rk9.png width=700>

https://vita.had.co.nz/papers/tidy-data.pdf
