---
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Packages.

```{r}
library(tidyverse)
```

Data.

```{r}
diamonds %>% 
  select(price, carat, x, y, z)
```

### Explore outliers analytically, and by zooming into a plot

Outliers:

* Unusual data.
* They don't seem to fit the pattern.
* May be data-entry errors.

Example:

* Why are the limits of the `y` axis so wide?

```{r}
ggplot(diamonds) + 
  geom_histogram(aes(x = y))
```

Let's zoom in! Can you now see what's going on?

```{r}
ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) +
  coord_cartesian(ylim = c(0, 50))
```

Here is another way to "zoom in". Why in this case it isn't a good way?

```{r}
ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) +
  ylim(c(0, 50))
```

Let's now explore those outliers analytically. Can you see why they are errors?

```{r}
unusual <- diamonds %>% 
  filter(y < 3 | y > 20) %>% 
  select(price, x, y, z) %>%
  arrange(y)
unusual
```

Tip: Repeat your analysis with and without the outliers:

* If they have substantial effect, only drop them if you can justify.

* If they have minimal effect, you can can replace them with missing values.

### Transforming incorrect values into missing values

How can you remove data-entry errors?

1. Remove the entire row. Why this may not be a good idea?

```{r}
diamonds2 <- diamonds %>% 
  filter(between(y, 3, 20))
```

2. Replace them with missing values.

```{r}
diamonds2 <- diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y))
```

Note the warning:

```{r}
ggplot(data = diamonds2, mapping = aes(x = x, y = y)) + 
  geom_point()
```

### Convert missing values into a logical variable so you can visualize them

Convert missing values into logical values, then visualize them.

```{r}
diamonds3 <- diamonds2 %>% 
  mutate(is_error = is.na(y))

ggplot(diamonds3) +
  geom_bar(aes(is_error))
```

Zoom in.

```{r}
diamonds3 %>% 
  count(is_error)

ggplot(diamonds3) +
  geom_bar(aes(is_error)) +
  coord_cartesian(ylim = c(0, 100))
```
