---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(tidyverse)
```

### Questions

* An exploratory Data Analysis (EDA) helps you understand your data.

* Use **questions** as tools to help you focus on a specific part of your data.

* Two types of questions are always useful:

    1. What type of **variation** occurs within my variables?
    
    2. What type of **covariation** occurs between my variables?

### Definitions

* Variable ~ column.

* Value ~ cell.

* Observation ~ row.

* Tabular data ~ spreadsheet.

* Tidy data ([paper](https://vita.had.co.nz/papers/tidy-data.pdf),
[tidyr](https://tidyr.tidyverse.org/) package).

> Tidy datasets are easy to manipulate, model and visualise, and have a specific structure: each variable is a column, each observation is a row.

<img src=https://i.imgur.com/nBC5Rk9.png width=700>

Example:

```{r}
diamonds
```

### Variation

Variation: The tendency of the values of a variable to change from measurement
to measurement.

* Every variable has its own pattern of variation. 

* Plotting the distribution of a variable helps you understand that pattern.

#### Exploring the distribution of categorical variables

A categorical variable can only take one of a small set of values.

```{r}
diamonds %>%
  relocate(cut)
```

Visualize:

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

By hand:

```{r}
diamonds %>% 
  count(cut)
```

A continuous variable can take any of an infinite set of ordered values.

```{r}
diamonds %>% 
  relocate(carat)
```

Visualize:

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)
```

By hand:

```{r}
diamonds %>% 
  count(cut_width(carat, 0.5))
```

* Tip: Different binwidths can reveal different patterns.

```{r}
smaller <- diamonds %>% 
  filter(carat < 3)
  
ggplot(data = smaller, mapping = aes(x = carat)) +
  geom_histogram(binwidth = 0.1)
```

* Overlay multiple histograms in the same plot:`geom_freqpoly()`.

```{r}
ggplot(data = smaller, mapping = aes(x = carat, colour = cut)) +
  geom_freqpoly(binwidth = 0.1)
```

### Resources

* [R for data science (book)](https://r4ds.had.co.nz/index.html).

* [ggplot2](https://ggplot2.tidyverse.org/):

    * [Template](https://r4ds.had.co.nz/data-visualisation.html#a-graphing-template).

    * [Reference](https://ggplot2.tidyverse.org/reference/index.html).

    * [Cheatsheet](https://ggplot2.tidyverse.org/index.html#cheatsheet).

* [dplyr](https://dplyr.tidyverse.org/):

    * [Reference](https://dplyr.tidyverse.org/reference/index.html).

