---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Packages.

```{r}
library(tidyverse)
```

Data.

`?diamonds`:

* `price`: price in US dollars.
* `carat`: weight of the diamond (0.2–5.01).
* `x, y, z`: length, width, depth in mm.

```{r}
diamonds %>% 
  select(price, carat, x, y, z)
```

An overview.

```{r}
diamonds %>% 
  glimpse()
```

### Visualize the most common and rare values in a dataset

What a bar chart or histogram can tell you:

* Tall bars show common values.

* Shorter bars show less-common values. 

* No bars reveal values that were not seen in your data.

General questions:

* Which values are the most common? Why?

* Which values are rare? Why? Does that match your expectations?

* Can you see any unusual patterns? What might explain them?

Specific questions for the plot below:

* Why are there more diamonds at whole carats and common fractions of carats?

* Why are there more diamonds slightly to the right of each peak than there are
slightly to the left of each peak?

```{r}
smaller <- diamonds %>% 
  filter(carat < 3)

ggplot(data = smaller, mapping = aes(x = carat)) +
  geom_histogram(binwidth = 0.01)
```

Clusters of similar values suggest that subgroups exist in your data. To understand the subgroups, ask:

* How are the observations within each cluster similar to each other?

* How are the observations in separate clusters different from each other?

* Why might the appearance of clusters be misleading? (Consider bin-width.)
