---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Defensive workflow

(Find RStudio commands with: Shift + Ctrl/Cmd + P)

[Start from a blank slate every time you restart R](https://rstats.wtf/save-source.html):

* Save workspace on quit: Never.
* Load workspace on start: Off.
* Restart R session.

[Use RStudio projects](https://rstats.wtf/project-oriented-workflow.html#rstudio-projects):

* Create a new project.
* Open project.
* File > Recent projects

Tips:

* Have a dedicated folder for your projects.
* Launch recent projects with: Alt + F > J
* Launch projects with a dedicated app (Alfred, Albert, etc.)

[Practice safe paths](https://rstats.wtf/safe-paths.html):

```{r error = TRUE}
# Good
path <- here::here("data", "greeting.txt")
path
readLines(path)

# Bad
path <- "/a/fragile path/that/only/i/have/data/greeting.txt"
readLines(path)
```

[How to name files](https://rstats.wtf/how-to-name-files.html):

<img src=https://i.imgur.com/Lo7JFxc.png width=700>

https://speakerdeck.com/jennybc/how-to-name-files?slide=3

### The data science workflow

<img src=https://i.imgur.com/MitpHrJ.png width=700>

https://r4ds.had.co.nz/introduction.html

### The data science toolkit

<img src=https://i.imgur.com/i61n3xb.png width=700>

https://www.tidyverse.org/

```{r packages}
library(tidyverse)
```

### Visualize: The ggplot template

<img src=https://i.imgur.com/WsoXgV2.png width=700>

Template:

```
    ggplot(data = <DATA>) +
      <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

Example:

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

https://ggplot2.tidyverse.org/

### Transform: Key dplyr verbs and how to compose them

Example:

```{r}
diamonds
```

```{r}
select(diamonds, cut, price)
```

Same:

```r
# Same
diamonds %>% select(cut, price)
```

```{r}
diamonds %>% 
  select(cut, price) %>% 
  count(cut) %>% 
  filter(n > 10000)
```

Same but less readable:

```r
filter(count(select(diamonds, cut, price), cut), n > 10000)
```

https://dplyr.tidyverse.org/

### Tidy data: What and why

> Tidy datasets are easy to manipulate, model and visualise, and have a specific structure: each variable is a column, each observation is a row.

<img src=https://i.imgur.com/nBC5Rk9.png width=700>

https://vita.had.co.nz/papers/tidy-data.pdf
